```{r}
library(dplyr)
library(tidyr)
library(readr)
library(glue)
library(ggplot2)
library(scales)
library(stringr)
```


# DPLYR and GGPLOT tutorial code

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("gapminder")

library(dplyr)
library(ggplot2)
library(gapminder)

# Preview the data
head(gapminder)

# Check structure
str(gapminder)
```


```{r}
# Filter for Slovenia only
gapminder  %>%  
  filter(country == "Slovenia")
```

```{r}
# Select just country, year, and life expectancy
gapminder %>% 
  select(country, year, lifeExp)
```

```{r}
# Add column GDP in billions
gapminder %>%
  mutate(gdp_billion = gdpPercap * pop / 1e9)

```

```{r}
# Average life expectancy per continent
gapminder %>%
  group_by(continent) %>%
  summarise(avg_lifeExp = mean(lifeExp))

```

```{r}
gapminder %>%
  filter(country == "Slovenia") %>%
  ggplot(aes(x = year, y = lifeExp)) +
  geom_line(color = "blue", size = 1.2) +
  labs(title = "Life Expectancy in Slovenia Over Time",
       x = "Year",
       y = "Life Expectancy") +
  theme_bw()


ggplot(gapminder, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() +
  scale_x_log10() + # log scale to make it more readable
  labs(title = "Life Expectancy vs GDP per Capita",
       x = "GDP per Capita (log scale)",
       y = "Life Expectancy") +
  theme_bw()


```





# App1 vs App2 retention data 
```{r}
# import the data
retention_df <- read_csv("analysis_retention.csv")

retention_plot_data <- retention_df %>% 
  mutate(daySinceInstall = as.numeric(activeDate - installDate)) %>% 
  group_by(app, installDate, daySinceInstall) %>% 
  summarise(activeUsers = n_distinct(user_label)) %>% 
  ungroup() %>%
  group_by(app, installDate) %>% 
  mutate(cohortUsers = max(activeUsers)) %>% 
  ungroup() %>%
  filter(daySinceInstall >= 0)   # Keep only valid daySinceInstall values

```

```{r}
retention_plot_data %>% 
  filter(daySinceInstall %in% c(1,3,7)) %>% 
  ggplot(aes(installDate, activeUsers/cohortUsers, color = app)) +
    geom_line() +
    geom_point() +
    facet_grid(daySinceInstall~.) +
    scale_y_continuous(labels = scales::percent) +
    labs(title = "Retention for two apps", x = "install date", y = "% active users") +
    expand_limits(y = 0) +
    theme_bw()
```

```{r}
retention_plot_per_app <- retention_plot_data %>% 
  group_by(app, daySinceInstall) %>% 
  summarise(cohortUsers = sum(cohortUsers), activeUsers = sum(activeUsers)) %>% 
  ungroup() %>%
  mutate(retention_rate = activeUsers/cohortUsers)
  
retention_plot_per_app %>% 
  ggplot(aes(daySinceInstall, retention_rate, color = app)) +
    geom_line() +
    geom_point() +
    scale_y_continuous(labels = scales::percent) +
    labs(title = "Retention for two apps", x = "install date", y = "% active users") +
    expand_limits(y = 0) +
    theme_bw()
```

```{r}
retention_plot_relative <- retention_plot_data %>% 
  group_by(app, daySinceInstall) %>% 
  summarise(cohortUsers = sum(cohortUsers), activeUsers = sum(activeUsers)) %>% 
  ungroup() %>%
  mutate(retention_rate = activeUsers/cohortUsers)

retention_app_1 <- retention_plot_relative %>% 
  filter(app == "app 1") %>% 
  rename(retention_rate_app_1 = retention_rate) %>% 
  select(retention_rate_app_1, daySinceInstall)

retention_plot_relative %>% 
  merge(retention_app_1, by = "daySinceInstall") %>% 
ggplot(aes(daySinceInstall, retention_rate/retention_rate_app_1, color = app)) +
    geom_line() +
    geom_point() +
    scale_y_continuous(labels = scales::percent) +
    labs(title = "Retention for two apps", x = "install date", y = "% active users") +
    expand_limits(y = 0) +
    theme_bw()
```



# MTA vs MTT data - feature usages
```{r}
# import data
features_mta <-  read_csv("features_mta.csv")
features_mtt <-  read_csv("features_mtt.csv")
day_users_mta <- read_csv("day_users_mta.csv")
day_users_mtt <- read_csv("day_users_mtt.csv")
```

## minigame-play
```{r}
# Filter for "minigame_play" feature and assign app labels
minigame_mta <- features_mta %>%
  filter(feature == "minigame_play") %>%
  mutate(app = "mta")

minigame_mtt <- features_mtt %>%
  filter(feature == "minigame_play") %>%
  mutate(app = "mtt")

# Combine both datasets
minigame_combined <- bind_rows(minigame_mta, minigame_mtt)

# Prepare daily user data with app labels
day_users <- bind_rows(
  day_users_mta,
  day_users_mtt
)

# Merge event data with daily user data
minigame_merged <- minigame_combined %>%
  left_join(day_users, by = c("app","daySinceInstall")) %>%
  mutate(avgCAU = actions / cohortUsers)  # Calculate average usage per CAU

# Create a well-presented ggplot visualization
ggplot(minigame_merged, aes(x = daySinceInstall, y = avgCAU, color = app)) +
  geom_line(size = 1.2) +  # Thicker line for better visibility
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = 0.05)) +
  facet_wrap(~feature, scales = "free_y") +  # Separate plots for each event type
  labs(
    title = "Feature Usage per Cohort Active User (CAU)",
    subtitle = "Comparison of 'minigame_play' events across MTA and MTT",
    x = "Days Since Install (DSI)",
    y = "Average Feature Usage per CAU",
    color = "Application"
  ) +
  theme_bw()

```


## For multiple features
```{r}
filtered_features_mta <- features_mta %>% 
  filter(!is.na(feature)) %>%
  mutate(app = "mta")

filtered_features_mtt <- features_mtt %>% 
  filter(!is.na(feature)) %>%
  mutate(app = "mtt")

features_df <- filtered_features_mta %>% 
  rbind(filtered_features_mtt) %>% 
  left_join(day_users) %>% 
  mutate(group_events = str_extract(feature, "[a-z]+")) %>% 
  mutate(avgCAU = actions/cohortUsers) %>% 
  split(.$group_events, drop = T)

features_df %>% 
    purrr::walk(function(data) {
    fplot <- ggplot(data, aes(x = daySinceInstall,  y = avgCAU, color = app)) + 
        geom_line() +
        scale_y_continuous(limits = c(0, NA)) +
        facet_wrap(feature~.)+
        labs(title = paste0("Number of feature usage per CAU: ", data$group_events[1]), x = "dsi", y = "Avg per CAU", color = "Event") +
      theme_bw()
    plot(fplot)
})

dataframe_features_CAU <- filtered_features_mta %>% 
  rbind(filtered_features_mtt) %>% 
  left_join(day_users) %>% 
  mutate(group_events = str_extract(feature, "[a-z]+")) %>% 
  mutate(cauPerc = users/cohortUsers) %>%
  split(.$group_events, drop = T)
  
dataframe_features_CAU %>% 
    purrr::walk(function(data) {
    fplot <- ggplot(data, aes(x = daySinceInstall,  y = cauPerc, color = app)) + 
        geom_line() +
        scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
        facet_wrap(feature~.)+
        labs(title = paste0("Feature usage: ", data$group_events[1]), x = "dsi", y = "% CAU", color = "Event") +
      theme_bw()
     plot(fplot)
  })
```




# AB TEST RESULTS
## Import data
```{r}
library(dplyr)
library(ggplot2)
library(readr)

ab_test_minigames_df <- read_csv("ab_test_percent_users_minigames.csv") %>% 
  mutate(groupDescription = case_when( # add a description for each group for easier understanding
    testGroup == "A" ~ "A: control group",
    testGroup == "B" ~ "B: 8 most popular minigames",
    testGroup == "C" ~ "C: 8 least popular minigames",
    testGroup == "D" ~ "D: 8 most similar to MTA",
    TRUE ~ "Unknown"))

ab_test_retention_df <- read_csv("ab_test_retention.csv") %>% 
  mutate(groupDescription = case_when(
    testGroup == "A" ~ "A: control group",
    testGroup == "B" ~ "B: 8 most popular minigames",
    testGroup == "C" ~ "C: 8 least popular minigames",
    testGroup == "D" ~ "D: 8 most similar to MTA",
    TRUE ~ "Unknown"))
```

```{r}
ab_test_minigames_df
```


## AB test retention
```{r fig.height=8, fig.width=20}
ab_test_retention_df %>% 
  group_by(testGroup) %>% 
  mutate(testUsers = max(ifelse(daySinceInstall == 0, users, 0))) %>% # calculate number of users per group
  ungroup() %>%
  mutate(retention = users/testUsers) %>% 
  
  ggplot(aes(x = daySinceInstall, y = retention, color = groupDescription)) +
    geom_line() +
    geom_point() +
    theme_bw() + 
    scale_y_continuous(labels = scales::percent) + # transforms 0.1 to 10%
    scale_x_continuous(minor_breaks = seq(0 , 15, 1), breaks = seq(0, 15, 1)) + # add vertical lines for each day
    scale_color_discrete(name = "Groups") + # rename the legend title
    theme(text = element_text(size = 20)) +
    expand_limits(y = 0) + # includes y = 0
    labs(title = "Retention")
```

## AB test zoomed in retention
```{r fig.height=8, fig.width=20}
ab_test_retention_df %>% 
  group_by(testGroup) %>% 
  mutate(testUsers = max(ifelse(daySinceInstall == 0, users, 0))) %>% 
  ungroup() %>%
  mutate(retention = users/testUsers) %>% 
  
  ggplot(aes(x = daySinceInstall, y = retention, color = groupDescription)) +
    geom_line() +
    geom_point() +
    theme_bw() + 
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(minor_breaks = seq(0 , 15, 1), breaks = seq(0, 15, 1)) +
    scale_color_discrete(name = "Groups") +
    coord_cartesian(y = c(0.1, 0.3)) +
    theme(text = element_text(size = 20)) +
    labs(title = "Zoomed-in retention")
```

## AB test relative retention
```{r fig.height=8, fig.width=20}
calculated_retention <- ab_test_retention_df %>% 
  group_by(testGroup, groupDescription) %>% 
  mutate(testUsers = max(ifelse(daySinceInstall == 0, users, 0))) %>% 
  ungroup() %>%
  mutate(retention = users/testUsers) %>% 
  select(daySinceInstall, testGroup, retention, groupDescription)

group_a <- calculated_retention %>% 
  filter(testGroup == "A") %>% 
  rename(controlGroup = testGroup, controlRetention = retention)

calculated_retention %>% 
  merge(group_a, by = "daySinceInstall") %>% 
  mutate(relativeDifference = retention / controlRetention - 1) %>% 
  
  ggplot(aes(x = daySinceInstall, y = relativeDifference, color = groupDescription.x)) +
    geom_line() +
    geom_point() +
    theme_bw() + 
    scale_y_continuous(labels = scales::percent) +
    scale_x_continuous(minor_breaks = seq(0 , 15, 1), breaks = seq(0, 15, 1)) +
    scale_color_discrete(name = "Groups") +
    expand_limits(y = 0) +
    theme(text = element_text(size = 20)) +
    labs(title= "Relative difference in retention between test groups and control group", y = "% difference from control")
```


## Percent of users playing minigames - bar
```{r fig.width=15, fig.height=8}
ab_test_minigames_df %>% 
  ggplot(aes(x = minigame, y = percentOfUsers, fill = minigame)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~groupDescription, nrow = 2, ) +
    labs(
      title = "Percent of users playing minigames",
      x = "",
      y = "% users",
    ) +
    theme_bw() +
    theme(text = element_text(size = 20), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position = "none")
```

```{r fig.width=15, fig.height=8}
ab_test_minigames_df %>% 
  filter(percentOfUsers > 5) %>%
  ggplot(aes(x = minigame, y = percentOfUsers, fill = minigame)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    facet_wrap(~groupDescription, nrow = 2, ) +
    labs(
      title = "Percent of users playing minigames",
      x = "",
      y = "% users",
    ) +
    theme_bw() +
    theme(text = element_text(size = 20), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position = "none")
```

## Percent of users playing minigames - scatter
```{r fig.width=15, fig.height=8}
ab_test_minigames_df %>% 
  filter(percentOfUsers > 5) %>% 
  ggplot(aes(x = "games", y = percentOfUsers, color = minigame)) +
    geom_text(aes(label = minigame), hjust = 0.5, vjust = 1.5, position = position_jitter(width = 0.5, height = 0), size = 7) +
    facet_wrap(~groupDescription, nrow = 1) +
    labs(
      title = "Percent of users playing minigames",
      x = "",
      y = "% users",
    ) +
    expand_limits(y = 0) +
    theme(text = element_text(size = 30), strip.text = element_text(size = 20)) +
    guides(color = "none") +
    theme_bw() 
```

## Selected minigames
```{r}
group_a_data <- ab_test_minigames_df %>% 
  filter(testGroup == "A")

group_b_minigames <- ab_test_minigames_df %>% 
  filter(percentOfUsers > 5, testGroup == "B") %>%
  distinct(minigame)

group_b_data <- group_a_data %>% 
  merge(group_b_minigames) %>% 
  mutate(testGroup = "B")
  
group_c_minigames <- ab_test_minigames_df %>% 
  filter(percentOfUsers > 5, testGroup == "C") %>%
  distinct(minigame)

group_c_data <- group_a_data %>% 
  merge(group_c_minigames) %>% 
  mutate(testGroup = "C")

group_d_minigames <- ab_test_minigames_df %>% 
  filter(percentOfUsers > 5, testGroup == "D") %>%
  distinct(minigame)

group_d_data <- group_a_data %>% 
  merge(group_d_minigames) %>% 
  mutate(testGroup = "D")

selected_games_df <- group_a_data %>% rbind(group_b_data, group_c_data, group_d_data)

selected_games_df
```

```{r fig.width=15, fig.height=8}
selected_games_df %>% 
  ggplot(aes(x = "games", y = percentOfUsers, color = minigame)) +
    geom_text(aes(label = minigame), hjust = 0.5, vjust = 1.5,position = position_jitter(width = 0.5, height = 0), size = 7) +
    facet_wrap(~testGroup, nrow = 1) +
    labs(
      title = "Percent of users playing minigames",
      x = "",
      y = "% users",
    ) +
    scale_y_continuous(limits = c(0, NA)) +
    theme(text = element_text(size = 30), strip.text = element_text(size = 20)) +
    guides(color = "none") +
    theme_bw() 
```

## Average number of minigames
```{r fig.width=15, fig.height=9}
ab_test_minigames_df %>% 
  filter(percentOfUsers > 5) %>% 
  ggplot(aes(x = minigame, y = minigamePlaysPerUser, fill = minigame)) +
    geom_bar(stat = "identity") +
    facet_wrap(~groupDescription, nrow = 2, ) +
    labs(
      title = "Average number of plays per user per minigame",
      x = "",
      y = "% users",
    ) +
    theme_bw() +
    theme(text = element_text(size = 20), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    theme(legend.position = "none")
```








# APPENDIX

## DAU
```{r}
# Simulating example data with a drop from Jan 15 to Jan 22
set.seed(123)
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  dau = 500 + (1:30 * 10) + sample(0:50, 30, replace = TRUE) 
)

# Introduce a drop from Jan 15 to Jan 22
df <- df %>%
  mutate(dau = ifelse(date >= as.Date("2024-01-15") & date <= as.Date("2024-01-22"), 
                      dau - sample(100:200, 1), dau))

# Plot using ggplot
ggplot(df, aes(x = date, y = dau)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(aes(color = ifelse(date >= as.Date("2024-01-15") & date <= as.Date("2024-01-22"), 
                                "Drop", "Normal")), size = 2) + 
  scale_color_manual(values = c("Drop" = "Black", "Normal" = "black")) +
  labs(title = "Daily Active Users (DAU) ",
       x = "Date",
       y = "Active Users",
       color = "DAU Status") +
  theme_bw() +
  scale_y_continuous(limits = c(0,NA))+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

## New installs
```{r}
# Simulate example data: 30 days of new installs
set.seed(123)
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  installs = round(runif(30, min = 50, max = 200)) # Random normal installs
)

# Introduce an exponential spike on January 15 that decays exponentially until Jan 20
spike_day <- as.Date("2024-01-15")
decay_days <- 5  # Duration of exponential drop (Jan 15 - Jan 20)
spike_value <- 500  # Peak installs on Jan 15
decay_rate <- 0.2  # Controls how fast it drops

df <- df %>%
  mutate(installs = ifelse(date == spike_day, spike_value, installs), # Spike on Jan 15
         installs = ifelse(date > spike_day & date <= spike_day + decay_days, 
                           spike_value * exp(-decay_rate * as.numeric(date - spike_day)), 
                           installs)) # Exponential drop

# Plot using ggplot2
ggplot(df, aes(x = date, y = installs)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(aes(color = ifelse(date >= spike_day & date <= spike_day + decay_days, 
                                "Spike & Drop", "Normal")), size = 2) + 
  scale_color_manual(values = c("Spike & Drop" = "black", "Normal" = "black"),guide = "none") +
  labs(title = "New Installs with Exponential Spike and Drop",
       x = "Date",
       y = "New Installs",
       color = "Install Pattern") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )+
  scale_y_continuous(limits = c(0,NA))
```

## Daily playtime users
```{r}
# Simulate daily playtime per user (in minutes)
set.seed(123)
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  playtime = round(runif(30, min = 30, max = 90)) # Normal daily playtime (random values)
)

# Introduce a small increase on January 15 that remains constant afterward
increase_day <- as.Date("2024-01-15")
increase_value <- 50  # Small increase in minutes

df <- df %>%
  mutate(playtime = ifelse(date >= increase_day, playtime + increase_value, playtime)) 

# Plot using ggplot2
ggplot(df, aes(x = date, y = playtime)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(color = "black", size = 2) + 
  labs(title = "Daily Playtime per User with a Small Increase",
       x = "Date",
       y = "Playtime (minutes)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits = c(0,NA))

```

## Daily sessions per user
```{r}
# Simulate daily sessions per user
set.seed(123)
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  sessions = round(runif(30, min = 2, max = 6)) # Random sessions per user
)

# Introduce a small increase on January 15 that remains constant afterward
increase_day <- as.Date("2024-01-15")
increase_value <- 1  # Small increase in sessions per user

df <- df %>%
  mutate(sessions = ifelse(date >= increase_day, sessions + increase_value, sessions)) 

# Plot using ggplot2
ggplot(df, aes(x = date, y = sessions)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(color = "black", size = 2) + 
  labs(title = "Daily Sessions Per User with a Small Increase",
       x = "Date",
       y = "Sessions per User") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_y_continuous(limits = c(0,NA))
```

## Average Session Duration
```{r}
# Set seed for reproducibility
set.seed(123)

# Generate random session durations 
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  session_duration = round(runif(30, min = 5, max = 15) + rnorm(30, mean = 0, sd = 2), 1) # Adding more randomness
)

# Ensure no negative values
df <- df %>%
  mutate(session_duration = ifelse(session_duration < 2, 2, session_duration)) # Set a lower bound

# Plot using ggplot2
ggplot(df, aes(x = date, y = session_duration)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(color = "black", size = 2) + 
  labs(title = "Average Session Duration Per User (Randomized)",
       x = "Date",
       y = "Session Duration (minutes)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    scale_y_continuous(limits = c(0,NA))



```

## Revenue
```{r}
# Simulate daily revenue per user 
set.seed(123)
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  revenue = round(runif(30, min = 0.15, max = 0.3), 3)  # Round to 3 decimal places
)

# Introduce a small increase on January 15 that remains constant afterward
increase_day <- as.Date("2024-01-15")
increase_value <- 0.5  # Small increase in revenue

df <- df %>%
  mutate(revenue = ifelse(date >= increase_day, revenue + increase_value, revenue)) 

# Plot using ggplot2
ggplot(df, aes(x = date, y = revenue)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(color = "black", size = 2) + 
  labs(title = "Revenue per Daily active User",
       x = "Date",
       y = "Revenue (USD)") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_y_continuous(limits = c(0, NA))

```

## Retention
```{r}
# Set seed for reproducibility
set.seed(123)

# Simulate improving retention rates over 30 days
df <- data.frame(
  date = seq(as.Date("2024-01-01"), by = "day", length.out = 30),
  D1 = round(seq(40, 45, length.out = 30) + rnorm(30, mean = 0, sd = 1), 1),  # Gradual increase for D1 retention
  D7 = round(seq(15, 20, length.out = 30) + rnorm(30, mean = 0, sd = 0.5), 1),  # Slight increase for D7 retention
  D30 = round(seq(5, 8, length.out = 30) + rnorm(30, mean = 0, sd = 0.3), 1)   # Slow but steady growth for D30 retention
)

# Reshape data for ggplot
df_long <- df %>%
  pivot_longer(cols = c(D1, D7, D30), names_to = "Retention_Day", values_to = "Retention_Rate")

# Plot using ggplot2
ggplot(df_long, aes(x = date, y = Retention_Rate, color = Retention_Day)) +
  geom_line(size = 1) + 
  geom_point(size = 2) + 
  scale_color_manual(values = c("D1" = "blue", "D7" = "red", "D30" = "green")) +
  labs(title = "User Retention Over Time (Improving Trend in D1, D7, D30)",
       x = "Date",
       y = "Retention Rate (%)",
       color = "Retention Day") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


